Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood

### 2021-06-30 doc/temp-sentiment.tex

%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
Higher temperatures associated with climate change are expected to have major impacts on human mental health. Indeed, previous work has found strong associations between heat and mental health outcomes. These studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race at coarse spatial scales. However, using expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status, we show that there are stark differences in the effects of heat depending on neighborhood characteristics. Here we show that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods. We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway. This paper presents the first evidence of heterogeneity in the vulnerability of mental health to heat.
\end{abstract}
\begin{document}

@@

\section*{Main}
%Overview
Climate change is impacting many aspects of human well-being. A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \cite{pachauri2014climate}. However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr, hayes_climate_2018}. While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}. Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, and linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}. Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of increased heat on mental health \cite{Berry2018Apr}, especially for identifying the most vulnerable communities.

%Theoretical: frameworks and pathways
Researchers have proposed a number pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}. Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul, sadiq_impact_2019}. Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007, WHO2007}. Additionally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.

%Empirical: Heat and mental health
Drawing on these frameworks, results from an emerging literature are beginning to document linkages between higher temperatures worsened mental health outcomes. Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}. Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}. Higher temperatures are also linked to increased mortality among people with pre-existing mental health disorders \cite{Hansen2008Oct}.  Finally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}, although these studies have relied on retrospective survey data and have not used data from across the diurnal cycle.

The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress. For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}. Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}. A major challenge for these studies, however, is their reliance on data that is typically aggregated to the municipality or county level. As a consequence, these studies are often restricted to between-county metrics of vulnerability and cannot take into account neighborhood effects.

These findings of uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat. For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}. Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities. For example, Black Americans face discrimination in employment \cite{Reid2003, Kang2016, Penner2008} and housing access \cite{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat. Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \cite{Banzhaf2019}. Notably, there is often an intersectionality between different forms of vulnerability and oppression, with income and race likely having a compounding effect on vulnerability to environmental risks \cite{Kuran2020Nov}. Heterogeneity in vulnerability is also more pronounced within city regions rather than between cities, particularly in the United States where there can be considerable regional variability \cite{Cardona2012Jan}. Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}. Thus, there are theoretical reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.

While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely proxy for mental health at extremely precise spatial and temporal scales. Many studies use the mood expressed in social media posts as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}. For example, studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder (PTSD) in individuals even before their formal diagnosis \cite{Reece2017Oct}. Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}. In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}. Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \cite{Dodds2011, Dzogang2017}. Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}. Thus, posts on Twitter provide an proxy for mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures. To explore how neighborhood conditions affect the relationship between heat and mental heath, we draw on a data set of 243 million geolocated tweets from across the continental United States covering the period 2009-2019, each paired with data on local weather and neighborhood conditions.

@@

\section*{Discussion}
% Benefits of Twitter data/overview of findings
We find that Twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution. We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between increasing temperatures and declining mood. Moreover, we were able to show heterogeneity in vulnerability by neighborhood characteristics, improving on previous findings that found no heterogeneity at the county level \cite{Burke2018Aug, Mullins2019Dec}.

% Compare weekly changes in mood to weekly changes in suicides
We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a higher temperature of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in mood of 0.01. This is similar to the degree of change in mood over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}. Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of mood and suicides at some time scales. Moreover, like mood, suicides are affected by higher temperatures. While the mood expressed on Twitter is only a proxy for mental health and not a direct measurement, the two phenomena have similar temporal patterns and respond similarly to heat. Thus, the changes in expressed mood we observe at neighborhood scales are very likely indicative of mental health outcomes such as suicides and hospitalizations.

% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
% Variations in vulnerability also indicates ADAPTATION
Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income. Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change. For example, Mullins et al. conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}. Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income. This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the poor and vulnerable. More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable. Additionally, our results suggest that reducing existing discrimination in housing and employment would contribute narrowing racial gaps in vulnerability and be an important:w
adaptation strategy.

Counter to previous studies, we find that the highest temperatures disproportionately affect poor and marginalized populations. Moreover, we find that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority. This is surprising, given that Hispanic peoples can be marginalized in income, language and housing. This may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans. It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis. Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health. Finally, while race is correlated with income in the US, we find that neighborhood racial characteristics more directly contributed to vulnerability to heat, independent of income. This suggests that racism and patterns of housing discrimination like the legacy of redlining contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups. Thus, our study shows that the degree of vulnerability to heat stress is less related to a single category or variable (e.g., race or income) but is more likely related to a variety of complex, intersecting dimensions like combinations of race and housing \cite{Kuran2020Nov}.

% Connect to bigger literature on vulnerability
This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts, as well as the importance of measuring impacts at fine spatial scales. Within in relatively wealthy nation of the United States, we find large heterogeneity in vulnerability. This adds to the growing literature showing that the impacts of climate change will be highly unequal and disproportionately borne by the poor and marginalized \cite{Thomas2019Mar}. Moreover, given that more impoverished neighborhoods in the United States see greater mental health impacts during heatwaves, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.

@@

% Caveats
% Mood is a fuzzy indicator but we have BIG DATA
While these findings are robust to different model specifications and metrics of mood, a limitation of data from Twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet. Moreover, while mood/sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed mood in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact. We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in mood in relation to the weather.

One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals. Moreover, in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are typically found that census block. For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day. Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American. Again, these issues mean we are likely under-estimating the true effect of heat on mental health for poor and Black people. A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}. Nevertheless, we found a large volume of tweets across all neighborhood types.

% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how highly unevenly these impacts will be distributed. People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}. While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed. Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect. By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States. These findings have significant implications for urban planning, climate and environmental justice, and mental health.

@@

We calculated the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet as temperature metrics that account humidity and other factors can better account for the impacts of heat stress on human health and well-being. The WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature due to evaporative cooling. The WBGT can be estimated given normal temperature, relative humidity, solar radiation, and wind speed. Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing \cite{budd2008wet}. Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \cite{rao2020projections}, increased crime \cite{hu2017impact}, increased mortality \cite{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \cite{vida2012relationship, ding2016importance}.

We derived relative humidity using methods described by Bolton et al. \cite{bolton_computation_1980} which use temperature, specific humidity, and pressure. We then calculated the WBGT using the formula described by Heo et al. \cite{heo2019comparison}.

Throughout the paper we give the median dry bulb temperature observed at each given wet bulb temp degree across our data set. It should be noted, however, that a wide range of dry bulb temperatures can be associated with a given wet bulb globe temperature, depending on humidity, wind speed, and sunlight.

@@

We present income and racial percentiles based on the observed percentiles across all tweets in our data set.

@@

In addition to our models exploring how income and race moderate the effect of heat on mood individually, we fit an additional model with an interaction terms for both race and income. Due to the complexity of a model with segmented terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine available to us. Thus, we used a continuous metric for both income as well as race (percent black). This model extends our main model to the following form:

\begin{equation}
 y = \begin{tabular}{l}
 $\beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) + 
 \beta_p p + \beta_{m_{r} p} m_{r} p + 
 \beta_{m_{i} p} m_{i} p + \beta_{m_{i} m_{r} p} m_{i} m_{r} p + $
 \\
 $\beta_s s + \beta_{m_{i} s} m_{i} s + 
 \beta_{m_{r} s} m_{r} s + \beta_{m_{i} m_{r} s} m_{i} m_{r} s + 
 \Phi + \epsilon$
 \end{tabular}
 \label{mod:race-inc}
\end{equation}

>>>>

%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
Higher temperatures associated with climate change are expected to have major impacts on human mental health, and previous work has found strong associations between heat and mental well-being. However, these studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race.  We therefore use expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status available at fine spatial and temporal scales, and find stark differences in the effects of heat depending on neighborhood characteristics. Here we show that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods. We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway. This paper presents the first evidence of neighborhood heterogeneity in the vulnerability of mental health to heat.
\end{abstract}
\begin{document}

@@

\section*{Main}
%Overview
Climate change is impacting many aspects of human well-being. A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \cite{pachauri2014climate}. However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr}. While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}. Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Schwartz2017Aug}, as well as linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Obradovich2018Oct}. Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of increased heat on mental health, especially for identifying the most vulnerable communities.

%Theoretical: frameworks and pathways
Researchers have proposed a number pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr}. Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul}. Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007}. Additionally, exposure to higher temperatures can affect productivity and income \cite{Burke2015Nov}, leading to second-order impacts on mental health outcomes.

%Empirical: Heat and mental health
Drawing on these frameworks, results from an emerging literature are beginning to document linkages between higher temperatures worsened mental health outcomes. Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Likhvar2011Jan}. Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}. Higher temperatures are also linked to increased mortality among people with pre-existing mental health disorders \cite{Hansen2008Oct}.  Finally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}, although these studies have relied on retrospective survey data and have not used data from across the diurnal cycle.

The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicates surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress. For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}. Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}. A major challenge for these studies, however, is their reliance on data that is typically aggregated to the municipality or county level. As a consequence, these studies are often restricted to between-county metrics of vulnerability and cannot take into account neighborhood effects.

These findings of uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat. For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}. Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities. For example, Black Americans face discrimination in employment \cite{Kang2016, Penner2008} and housing access \cite{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat. Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \cite{Banzhaf2019}. Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}. Thus, there are theoretical reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.

While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely proxy for mental health at extremely precise spatial and temporal scales. Many studies use the mood expressed in social media posts as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}. For example, studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder (PTSD) in individuals even before their formal diagnosis \cite{Reece2017Oct}. Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}. In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}. Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \cite{Dodds2011}. Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}. Thus, posts on Twitter provide an proxy for mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures. To explore how neighborhood conditions affect the relationship between heat and mental heath, we draw on a data set of 243 million geolocated tweets from across the continental United States covering the period 2009-2019, each paired with data on local weather and neighborhood conditions.

@@

\section*{Discussion}
% Compare weekly changes in mood to weekly changes in suicides
We found that the change from an optimum temperature to a higher temperature is associated with an overall decrease in mood similar in magnitude to the degree of change in mood from Monday to Saturday, a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}, and that the same level of heat in a poor or Black neighborhood is associated with a change in mood nearly matching the impact of a major Hurricane, an event associated with suicides and PTSD \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.  Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of expressed mood and mental health at specific time scales and following certain events.  This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations.

% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
% Variations in vulnerability also indicates ADAPTATION
Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income. Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change. For example, Mullins et al. conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}. Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income. This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the vulnerable and marginalized. More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable.

Examining the role of race in vulnerability, we find that majority Black non-Hispanic neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority. Surprisingly, we did not find an effect for people with Hispanic ethnicity, given that Hispanic peoples can be marginalized in employment and housing. These findings may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans. It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis. Additionally, there are many other marginalized groups in the United States, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health. Examining the effects of both race and income, we find that race is more important for vulnerability than income. This probably due in part to the fact that Black Americans have historically been forced through redlining into neighborhoods with fewer trees that are much more impacted by heat \cite{Hoffman2020, Locke2021}.  This suggests that racism and patterns of housing discrimination contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups, and that reducing existing discrimination in housing and employment is an important adaptation strategy.

% Connect to bigger literature on vulnerability
This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts.  Additionally, it shows how measuring outcomes at fine spatial scales necessary for understanding how factors like race and income matter for vulnerability \cite{Tong2021}. Even within a relatively wealthy nation, we find that high temperatures affect certain populations to the same degree as a major hurricane. Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.

@@

% Caveats
% Mood is a fuzzy indicator but we have BIG DATA
While these findings are robust to different model specifications and metrics of mood, there are important considerations in this analysis.  One caveat associated with this data is that we were only able to locate the tweets within the census block from which the tweet was sent, and in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are found that census block. For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day. These issue may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people. Another issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}. Nevertheless, we found a large volume of tweets across all neighborhood types.

% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how unevenly these impacts will be distributed. People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}. While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed. Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect. By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States. These findings have significant implications for urban planning, climate and environmental justice, and mental health.

@@

We calculated the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet. The WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature due to evaporative cooling. The WBGT can be estimated given normal temperature, relative humidity, solar radiation, and wind speed. Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing \cite{budd2008wet}. Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \cite{rao2020projections}, increased crime \cite{hu2017impact}, increased mortality \cite{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \cite{vida2012relationship, ding2016importance}.

We derived relative humidity using methods described by Bolton et al. \cite{bolton_computation_1980} which use temperature, specific humidity, and pressure. We then calculated the WBGT using the formula described by Heo et al. \cite{heo2019comparison}.  Throughout the paper we give the median dry bulb temperature observed at each given wet bulb temp degree across our data set. It should be noted, however, that a wide range of dry bulb temperatures can be associated with a given wet bulb globe temperature, depending on humidity, wind speed, and sunlight.

@@

We present income and percentiles based on the observed percentiles across all tweets in our data set.

@@

In addition to our models exploring how income and race moderate the effect of heat on mood individually, we fit an additional model with an interaction terms for both race and income. Due to the complexity of a model with segmented terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine available to us. Thus, we used a continuous metric for both income as well as race (percent black). This model extends our main model to the following form, interacting income and race with a temperature segmented spline as well as precipitation in sunshine:

\begin{equation}
 y = \beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) +  \beta_p p + \beta_{m_{r} p} m_{r} p + \beta_{m_{i} p} m_{i} p + \beta_s s + \beta_{m_{i} s} m_{i} s +  \beta_{m_{r} s} m_{r} s + 
 \Phi + \epsilon
 \label{mod:race-inc}
\end{equation}
